Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-02-18 ms/SI.rmd
Add new SI table for linear model fits
Rename 'model-fit' SI caption to 'gam-fit'

# Load analysis results
loadd(list = c(
  "taxonomic_data_for_si",
  "trait_data_for_si",
  "climate_ancova_results",
  "phylosig_binary_traits_strict",
  "correlated_evo_test_strict",
  "best_fit_div_models_moran",
  "func_comm_struc_combined"), 
  cache = canopy_cache)

@@

# Write out table with metadata
write_csv_with_meta(
  taxonomic_data_for_si, 
  here::here("results/Table_S1.csv"), 
  header,
  footer)
```

@@

# Write out table with metadata
write_csv_with_meta(
  trait_data_for_si, 
  here::here("results/Table_S2.csv"), 
  header,
  footer)
```

@@

# Write out table with metadata
write_csv_with_meta(
  climate_ancova_results_formatted, 
  here::here("results/Table_S3.csv"), 
  header,
  footer)
```

@@

# Write out table with metadata
write_csv_with_meta(
  phylosig_binary_traits_strict_formatted, 
  here::here("results/Table_S4.csv"), 
  header,
  footer)
```

@@

# Write out table with metadata
write_csv_with_meta(
  correlated_evo_test_strict_formatted, 
  here::here("results/Table_S5.csv"), 
  header,
  footer)
```

**`r s_table("model-fit")`** `r s_table_cap("model-fit")` (see separate file)

@@

```{r model-fit}
# Format table
best_fit_div_models_moran_formatted <-
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  )

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("model-fit") %>% wrap_quotes
)

@@

# Write out table with metadata
write_csv_with_meta(
  best_fit_div_models_moran_formatted, 
  here::here("results/Table_S6.csv"), 
  header,
  footer)

```

@@

```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
loadd(cwsd_scatterplot, cache = canopy_cache)

cwsd_scatterplot
```

>>>>

# Load analysis results
loadd(list = c(
  "taxonomic_data_for_si",
  "trait_data_for_si",
  "climate_ancova_results",
  "phylosig_binary_traits_strict",
  "correlated_evo_test_strict",
  "best_fit_div_models_moran",
  "func_comm_struc_combined",
  "climate_model_summaries",
  "climate_models",
  "div_el_models",
  "div_el_model_summaries",
  "cwsd_scatterplot"), 
  cache = canopy_cache)

@@

# Write out table with metadata
filename <- s_table("species-list") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

write_csv_with_meta(
  taxonomic_data_for_si, 
  here::here("results", filename), 
  header,
  footer)
```

@@

# Write out table with metadata
filename <- s_table("trait-values") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

write_csv_with_meta(
  trait_data_for_si, 
  here::here("results", filename), 
  header,
  footer)
```

@@

# Write out table with metadata
filename <- s_table("ancova") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

write_csv_with_meta(
  climate_ancova_results_formatted, 
  here::here("results", filename), 
  header,
  footer)
```

**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")` (see separate file)

MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods. 'NA' indicates not applicable.

```{r linear-fit}
# Combine summaries of models for diversity metrics
div_summaries <-
  div_el_model_summaries %>%
  left_join(
    select(div_el_models,
           var, model_type, delta_AICc, morans_I, I_pval), 
    by = c("var", "model_type")
  )

# Combine summaries of models for climate metrics
climate_summaries <-
  climate_model_summaries %>%
  left_join(
    select(climate_models,
           var, model_type, delta_AICc, morans_I, I_pval), 
    by = c("var", "model_type")
  )

# Combine summaries of diversity and climate metrics, reformat model names
table <-
  bind_rows(div_summaries, climate_summaries) %>%
  mutate(
    model_type = model_type %>% 
      str_replace_all("el_only", "Elevation") %>%
      str_replace_all("both", "Elevation + Growth habit") %>%
      str_replace_all("habit_only", "Growth habit") %>%
      str_replace_all("interaction", "Elevation + Growth habit + Elevation x Growth habit")
    )

# Reformat variable names and put in preferred order by joining
linear_fit_table <-
tribble(
  ~var, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "vpd_mean", "Mean VPD",
  "vpd_min", "Min. VPD",
  "temp_sd", "SD temp.",
  "vpd_sd", "SD VPD",
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA",
  "width", "Frond width"
) %>%
  left_join(table, by = "var") %>%
  select(
    Response,
    Model = model_type,
    AICc,
    f = statistic,
    `Adjusted R2` = adj.r.squared,
    P = p.value,
    `d.f.` = df,
    delta_AICc,
    `Moran's I` = morans_I,
    P_moran = I_pval
  )

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("linear-fit") %>% wrap_quotes
)

footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods. 'NA' indicates not applicable." %>% wrap_quotes

filename <- s_table("linear-fit") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

# Write out table with metadata
write_csv_with_meta(
  linear_fit_table, 
  here::here("results", filename), 
  header,
  footer)
```

@@

# Write out table with metadata
filename <- s_table("phylosig-binary-strict") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

# Write out table with metadata
write_csv_with_meta(
  phylosig_binary_traits_strict_formatted, 
  here::here("results", filename), 
  header,
  footer)
```

@@

# Write out table with metadata
filename <- s_table("corr-evo-strict") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

# Write out table with metadata
write_csv_with_meta(
  correlated_evo_test_strict_formatted, 
  here::here("results", filename), 
  header,
  footer)
```

**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")` (see separate file)

@@

```{r gam-fit}
# Format table
best_fit_div_models_moran_formatted <-
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  )

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("gam-fit") %>% wrap_quotes
)

@@

# Write out table with metadata
filename <- s_table("gam-fit") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

# Write out table with metadata
write_csv_with_meta(
  best_fit_div_models_moran_formatted, 
  here::here("results", filename), 
  header,
  footer)
```

@@

```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
cwsd_scatterplot
```

### 2020-02-18 ms/manuscript.Rmd
Add new SI table for linear model fits
Rename 'model-fit' SI caption to 'gam-fit'

Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r s_table("ancova")`; `r figure("climate")`).
Variables related to VPD differed by growth habit and elevation.
Mean and maximum or minimum temperature and VPD all decreased with elevation, whereas SD of temperature increased, but SD of VPD decreased with elevation.
Epiphytic dataloggers had higher values than terrestrial ones for variables related to VPD.

@@

In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r figure("heatmap")`).
Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("model-fit")`).
Spatial autocorrelation was not detected for any of the diversity metrics (`r s_table("model-fit")`).

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval{ }`r get_mod_result(model_summ_print, "sla", "p")`).
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).

We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).

@@

@@

**`r s_table("model-fit")`** `r s_table_cap("model-fit")`

>>>>

Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r s_table("ancova")`; `r figure("climate")`).
Variables related to VPD differed by growth habit and elevation.
Mean and maximum or minimum temperature and VPD all decreased with elevation, whereas SD of temperature increased, but SD of VPD decreased with elevation.
Epiphytic dataloggers had higher values than terrestrial ones for variables related to VPD.
Spatial autocorrelation was not detected in any of the climate models (`r s_table("linear-fit")`).

@@

In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r figure("heatmap")`).
Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("gam-fit")`).
Spatial autocorrelation was not detected in any of the best-fitting full-subsets models (`r s_table("gam-fit")`).

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities.
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).

We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
Spatial autocorrelation was not detected in any of the linear models for the community diversity metrics (`r s_table("linear-fit")`).

@@

**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")`

@@

**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")`
